\chapter*{Abstract}

This undergraduate thesis presents the design, implementation, and evaluation of a reproducible experimental pipeline based on Multiple Instance Learning (MIL) for weakly supervised classification of Whole Slide Images (WSI) in prostate cancer. The study addresses two central challenges in digital histopathology: interobserver variability in tumor grading and the limited availability of detailed patch-level annotations.\newline

The proposed methodology integrates a preprocessing stage focused on relevant tissue region selection using HSV color space filtering, deep feature extraction through a frozen ResNet-50 backbone, and the training of MIL architectures employing both classical pooling and attention-based aggregation mechanisms. The experimental framework was developed using the public SICAPv2 dataset and evaluated through strict patient-level cross-validation (\textit{GroupKFold}), ensuring complete separation between training and testing WSIs.\newline

The results demonstrate that attention-based MIL models, particularly SmABMIL, consistently outperform traditional pooling strategies in clinically relevant slide-level metrics. On average, the models achieved an F1-score above 0.83 and an AUC-ROC close to 0.86, along with improved inter-fold stability and a better balance between sensitivity and specificity. Furthermore, the incorporation of attention mechanisms enabled the generation of spatial relevance maps aligned with histopathological patterns associated with malignancy, enhancing the interpretability of the system.\newline

Overall, this work validates the feasibility of weakly supervised learning for prostate cancer classification in WSI and establishes a solid, reproducible methodological foundation for the development of clinically oriented decision-support systems in digital pathology.

\medskip
\textbf{Keywords:} Multi-instance learning (MIL); Whole Slide Images (WSI); Digital histopathology; Prostate cancer; Convolutional neural networks; Interpretability.